---
title: "More than Fortnite: An Analysis of Video Game Sales Using the *Dataset Video Game Sales with Ratings* by Rush Kirub"
subtitle: "Section: 1:20PM"
author: "Sandy Rosas, Pete Schultz & Tanush Samson"
date: "11/16/18"
output: 
  html_document: 
    fig_caption: yes
---


```{r,include=FALSE}
# These commands should always be near the top to ensure the libraries are loaded before the analysis.
library(mosaic)
library(Lock5withR)
library(readr)
library(dplyr)
#Games2 = read.csv(file = "VGS.csv")
Games2 = read.csv(file="C:\\Users\\tanus\\Documents\\VGS.csv")
head(Games2)
```

### Background
The videogame industry is a multi-billion dollar industry and a significant part of the world economy. We investigated certain trends and patterns within this industry to better understand it. The specific relationships we were interested in are Critic Score and Global Sales, Video Game Sales (by region) and civilian population, and Console and Global Sales. From this we are able to understand if games with a higher Critic Score sell more units globally, learn if the Japanese, European, and North American video game markets are representative of the amount of people living in each region, and to see how well a video game sells on PS4, Xbox One, and WiiU.


### Methods
We obtained our data from an online dataset called "Video Game Sales with Ratings" by Rush Kirubi from www.kaggle.com.  

With this dataset, we create a randomization distribution of the correlation between global sales (in millions of units) and critic scores. From this we check the probabilty of obtaining a sample correlation statistic as extreme as ours assuming there is no correlation.  

Further, we run a chi-square test to see if the proportion of video game sales in Japan, the European Union and North America match the proportion of civilian population in each region (proportion relative to the total population of all three regions combined).    

Lastly, we use a 90% confidence interval to find the average global sale (in millions of units) of a video game on a variety of consoles (PS4, Xbox One and WiiU).  

### Results

**RELATIONSHIP 1**  

$H_0: \rho = 0$  
$H_a: \rho > 0$  
where $\rho$ is the true population correlation between Global Sales (in millions of units) and Critic Score for all video games.

We obtain sample correlation of r = 0.245 with a p-value of essentially 0. This means that assuming the null hypothesis is true (that there is no correlation between global sales and critic score), there is essentially a 0% chance that we would obtain a sample correlation statistic as extreme as 0.245. Therefore, at the $\alpha$ = 0.01 level of significance, we reject the null hypothesis and conclude that there is sufficient evidence that supports the claim that there is a positive correlation between Global Sales and Critic Scores for all video games.  

```{r}
Games2.cor = filter(Games2,Critic_Score != "NA")
cor(Global_Sales~Critic_Score, data = Games2.cor)
set.seed(50000)
RandomizationCorrelation = do(1000)* cor(Global_Sales~shuffle(Critic_Score), data = Games2.cor)
gf_histogram(~cor, data = RandomizationCorrelation)
```
```{r}
prop(~cor >= 0.24547, data = RandomizationCorrelation)
```

**RELATIONSHIP 2**
 
An in-depth look into the distribution of global sales for the 3 main platforms: PS4, xbox One and WiiU.

```{r}
#Sales for PS4
gf_boxplot(Global_Sales ~ Platform, data = filter(Games2, Platform == "PS4"))
gf_boxplot(log(Global_Sales) ~ Platform, data = filter(Games2, Platform == "PS4"))
favstats(~Global_Sales, data = filter(Games2, Platform == "PS4"))
cdata(~Global_Sales, p = 0.9, data = filter(Games2, Platform == "PS4"))
```

```{r}
#Sales for Xbox One
gf_boxplot(Global_Sales ~ Platform, data = filter(Games2, Platform == "XOne"))
gf_boxplot(log(Global_Sales) ~ Platform, data = filter(Games2, Platform == "XOne"))
favstats(~Global_Sales, data = filter(Games2, Platform == "XOne"))
cdata(~Global_Sales, p = 0.9, data = filter(Games2, Platform == "XOne"))
```

```{r}
#Sales for WiiU
gf_boxplot(Global_Sales ~ Platform, data = filter(Games2, Platform == "WiiU"))
gf_boxplot(log(Global_Sales) ~ Platform, data = filter(Games2, Platform == "WiiU"))
favstats(~Global_Sales, data = filter(Games2, Platform == "WiiU"))
cdata(~Global_Sales, p = 0.9, data = filter(Games2, Platform == "WiiU"))
```


|         Console         |       PS4       |    Xbox One    |      Wii U      |
|:-----------------------:|:---------------:|:--------------:|:---------------:|
| 90% Confidence Interval (millions of units) | (0.020 , 3.908) | (0.10 , 2.872) | (0.020 , 2.105) |
|        Min Value  (millions of units)      |       0.01      |      0.01      |       0.01      |
|        Max Value   (millions of units)     |      14.63      |      7.39      |       7.09      |   

We are 90% confident that the true mean global sale of a video game on PS4 is between 0.020 and 3.908 million USD.  
We are 90% confident that the true mean global sale of a video game on Xbox One is between 0.010 and 2.872 million USD.  
We are 90% confident that the true mean global sale of a video game on WiiU is between 0.020 and 2.105 million USD.  


**RELATIONSHIP 3**

We expect the proportion of average video games sales in North America, Europeean Union countries and Japan to match the proportion of the respective sales in those regions.

$H_0: p_{NA} = 0.477, p_{EU} = 0.420, p_{JP} = 0.104$ 
$H_a:$ Some $p_i$ are not as specified.

|       | Population (Millions) | Population Proportion | Sales (Millions) | Sales Proportion |
|:-----:|:---------------------:|:---------------------:|:----------------:|:----------------:|
|   NA  |         582.4         |         0.4767        |       0.263      |      0.5422      |
|   EU  |         512.6         |         0.4195        |       0.145      |      0.2994      |
|   JP  |         126.8         |         0.1038        |       0.078      |      0.1620      |
| Total |         1221.8        |           1           |       0.486      |         1        |

```{r}
#Population for NA, EU and JP
pop = c(0.4767,0.4195,0.1038)
#Average sales for NA, EU and JP
Sales = c(263000,145000,78000)
#chi-squared test for significance 
xchisq.test(Sales, p = pop)
```




### Discussion
