---
title: "Math 260 Project Data Description"
subtitle: "Section: 1:20PM"
author: "Sandy Rosas, Pete Schultz & Tanush Samson"
date: "11/5/18"
output: 
  html_document: 
    fig_caption: yes
---


```{r,include=FALSE}
# These commands should always be near the top to ensure the libraries are loaded before the analysis.
library(mosaic)
library(Lock5withR)
library(readr)
library(dplyr)
#Games2 = read.csv(file = "VGS.csv")
Games2 = read.csv(file="C:\\Users\\tanus\\Documents\\VGS.csv")
head(Games2)
```

### Background
The videogame industry is a multi-billion dollar industry and a significant part of the world economy. We investigated certain trends and patterns within this industry to better understand it. The specific relationships we were interested in are American and Japanese sales, and console and genre. On top of this, we want to estimate the average number of global sales a video game has. From this we are able to better understand how large of an impact video games has on the global economy, learn which economy has a larger market for video games between the Americas and Japan, and to see if specific consoles favor specific genres of games more so than others .


### Methods
We obtained our data from an online dataset called "Video Game Sales with Ratings" by Rush Kirubi from www.kaggle.com.  

With this dataset, we bootstrap the correlation between global sales (in millions of units) and critic scores. From this we check the probabilty of obtaining a sample correlation statistic as extreme as ours assuming there is no correlation.  

Further, we run a two-sample t-test to see if the average Japanese sale of a video game  (in millions of units) differs from the average North American sale of a video game (in millions units).  

Lastly, we use a 90% confidence interval to find the average global sale (in millions of units) of video games on a variety of consoles (PS4, xbox One and WiiU).  

### Results

**RELATIONSHIP 1**  

$H_0: \rho = 0$  
$H_a: \rho > 0$  
where $\rho$ is the true population correlation between Global Sales (in millions of units) and Critic Score for all video games.

We obtain sample correlation of r = 0.245 with a p-value of essentially 0. This means that assuming the null hypothesis is true (that there is no correlation between global sales and critic score), there is essentially a 0% chance that we would obtain a sample correlation statistic as extreme as 0.245. Therefore, at the $\alpha$ = 0.01 level of significance, we reject the null hypothesis and conclude that there is sufficient evidence that supports the claim that there is a positive correlation between Global Sales and Critic Scores for all video games.  

```{r}
Games2.cor = filter(Games2,Critic_Score != "NA")
cor(Global_Sales~Critic_Score, data = Games2.cor)
set.seed(50000)
RandomizationCorrelation = do(1000)* cor(Global_Sales~shuffle(Critic_Score), data = Games2.cor)
gf_histogram(~cor, data = RandomizationCorrelation)
```
```{r}
prop(~cor >= 0.24547, data = RandomizationCorrelation)
```

**RELATIONSHIP 2**

An in-depth look into the distribution of global sales for the 3 main platforms: PS4, xbox One and WiiU.

```{r}
#Sales for PS4
gf_boxplot(Global_Sales ~ Platform, data = filter(Games2, Platform == "PS4"))
gf_boxplot(log(Global_Sales) ~ Platform, data = filter(Games2, Platform == "PS4"))
favstats(~Global_Sales, data = filter(Games2, Platform == "PS4"))
cdata(~Global_Sales, p = 0.9, data = filter(Games2, Platform == "PS4"))
```

```{r}
#Sales for Xbox One
gf_boxplot(Global_Sales ~ Platform, data = filter(Games2, Platform == "XOne"))
gf_boxplot(log(Global_Sales) ~ Platform, data = filter(Games2, Platform == "XOne"))
favstats(~Global_Sales, data = filter(Games2, Platform == "XOne"))
cdata(~Global_Sales, p = 0.9, data = filter(Games2, Platform == "XOne"))
```

```{r}
#Sales for WiiU
gf_boxplot(Global_Sales ~ Platform, data = filter(Games2, Platform == "WiiU"))
gf_boxplot(log(Global_Sales) ~ Platform, data = filter(Games2, Platform == "WiiU"))
favstats(~Global_Sales, data = filter(Games2, Platform == "WiiU"))
cdata(~Global_Sales, p = 0.9, data = filter(Games2, Platform == "WiiU"))
```


|         Console         |       PS4       |    Xbox One    |      Wii U      |
|:-----------------------:|:---------------:|:--------------:|:---------------:|
| 90% Confidence Interval | (0.020 , 3.908) | (0.10 , 2.872) | (0.020 , 2.105) |
|        Min Value        |       0.01      |      0.01      |       0.01      |
|        Max Value        |      14.63      |      7.39      |       7.09      | 

**RELATIONSHIP 3**

We expect the proportion of average video games sales in North America, Europeean Union countries and Japan to match the proportion of the respective sales in those regions.

$H_0: p_{NA} = 0.477, p_{EU} = 0.420, p_{JP} = 0.104$ 
$H_a:$ Some $p_i$ are not as specified.

|       | Population (Millions) | Population Proportion | Sales (Millions) | Sales Proportion |
|:-----:|:---------------------:|:---------------------:|:----------------:|:----------------:|
|   NA  |         582.4         |         0.4767        |       0.263      |      0.5422      |
|   EU  |         512.6         |         0.4195        |       0.145      |      0.2994      |
|   JP  |         126.8         |         0.1038        |       0.078      |      0.1620      |
| Total |         1221.8        |           1           |       0.486      |         1        |

```{r}
#Population for NA, EU and JP
pop = c(0.4767,0.4195,0.1038)
#Average sales for NA, EU and JP
Sales = c(263,145,78)
#chi-squared test for significance 
xchisq.test(Sales, p = pop)
```




### Discussion
